On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-03-10 Manuscript/manuscript.tex
comments/edits on simulation and conclusion

The Base R stats package has a built in function arima.sim that allows one to
simulate an AR(1) process, for which one can set the theoretical $\phi$
\citep{R}. By default, the mean of the process $\mu$ is 0 and the standard
deviation of the error term $\sigma_{\epsilon}$ is 1. For the purposes of this
study, the standard deviation of the process $\sigma_{x}$ should be controlled
at 1. The observations can be multiplied by $\sqrt{1 - \phi^2}$ so that the new
theoretical $\sigma_{x}$ is equivalent to 1, as shown below:
\begin{align}
X_{t} &= \epsilon_{t} + \phi X_{t-1};\\
Var(X_{t}) &= Var(\epsilon_{t} + \phi X_{t-1});\\
\sigma^2_{x} &= \sigma^2_{\epsilon} + \phi^2 \sigma^2_{x};\\
\sigma^2_{x}(1 - \phi^2) &= 1;\\
\sigma_{x}\sqrt{1 - \phi^2} &= 1.
\end{align}
The experimental factor of this study was $\phi$, as the temporal dependence
affects the performance of block bootstrap.
$\phi \in \{-0.4 -0.2, 0, 0.2, 0.4\}$.

As described in the last section, the block bootstrap method can be used to
estimate a parameter $\theta$ of an AR(1) process. This method can be
simulated with the tsboot function from the R boot package \citep{boot}
\citep{davison1997bootstrap}. In this study, $\theta$ is composed of the target parameters
$\mu = 0$, $\sigma_{x}$, and $\phi$. The sample mean $\bar{x}$ can be used as
the statistic to estimate $\mu$, the sample standard deviation $S_x$ can be
used as the statistic to estimate $\sigma$, and the autocorrelation function
of the sample at lag 1 $\rho_1$ can be used as the statistic to estimate
$\phi$. In the R implementation, these three statistics are incorporated into
a function that will apply each statistic to the original sample and to each
bootstrapped pseudo-sample. From a sample of length $n$, three types of block
bootstrap 95\% confidence intervals can be constructed for each target
parameter:
Percentile, $\hat{\theta}_{n}$-Centered,
and $BC_a$.

@@

@@

\bibliographystyle{chicago}
\bibliography{citations}[tp]

>>>>

We designed a simulation study to compare the performance of the different block
bootstrap confidence intervals. In particular, we generated time series $X_t$
from a 1st order auto-regressive (AR1) process:
\[
X_t = \phi X_{t-1} + \epsilon_t,
\]
where $\phi$ is an auto-regressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$.
The series $X_t$ has mean zero and variance
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$, we
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x^2 = 1$, the variance of $X_t$; and
2) $\phi$, the lag-1 auto-correlation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, \ldots, 1000\}$. In each configuration, we genrated 1000
replicates. For each replicate, we constructed three 95\% block bootstrap
confidence intervals for each parameter as described in the last section, and
estimated their actual coverage rates along with their 95\% confidence
intervals from the 1000 replicates. The block bootstrap sampling step was done
with function \texttt{tsboot} from R package \textsl{boot} \citep{boot}, with
block size \jy{size? also it was overlapping blocks. Describe the details here.}
The results are graphically displayed in Figures~\ref{fig:mu}--\ref{fig:phi}.

\jy{One paragraph for each parameter; write down your observations in bullet
  points and then organize them into a logically flowing paragraph.}

\jy{One summary paragraph at the end.}

@@

\jy{Discussion points:
  1) common practice on block bootstrap even for simple parameter such as the
  mean;
  2) further investigation needed for phi;
  3) impact on applied statsitics courses; student reactions (Dr. Schifano).
}

@@

\bibliographystyle{chicago}
\bibliography{citations}
